Fixed-Parameter Tractable Sampling for RNA Design with Multiple Target Structures

### 2018-05-24 RNARedPrint-Bioinfo2018.tex
changed citation

%To reduce the cost of the precomputation, we used a stacking energy model
%as a proxy for the Turner model. Indeed, we observed a strong linear
%dependency between the Turner energy and the
%emphasize that only the simple stacking energies are directly targeted
%(at $\pm$10\% tolerance, except for a few hard instances). Due to our
%linear adjustment, we finally achieve well-controlled energies in the
%realistic Turner energy model.
\begin{figure*}[t]
  \begin{center}
    %{\sf \bfseries A}\
    \includegraphicstop[width=0.99\textwidth]{Figs/energy_shift.pdf}\hfill
    %{\sf \bfseries B}\ \includegraphicstop[width=0.55\textwidth]{Figs/PKB00211_PKB00239_0_energy_distribution}
  \end{center}
  \caption{%
    Targeting specific energies for pseudoknotted structures using
    multi-dimensional Boltzmann sampling.  A) Linear fits between the
    energies in the stacking model to the realistic pseudoknot energy
    model by Dirks and Pierce (D\&P) for sampled sequences; the good
    match enables more efficient targeting of Turner energies based on
    targeting stacking model energies.  We show the fits for both
    target structures S1 and S2 of subfigure B.  B) D\&P energy
    distributions for two target structures S1 and S2. We show the
    results of targeting the respective free energies $(-30,-20)$,
    $(-30,-30)$, $(-25,-25)$, and $(-20,-30)$ for the target
    structures (annotated in red, yellow, and blue
    resepectively)---this demonstrates the effectivity of our adaptive
    multi-dimensional Boltzmann sampling procedure; moreover, for
    comparison, the distributions for uniform and Boltzmann
    samples---respectively associated with homogenous weights $1$ and
    $e^\beta$.
%    ($\delta=0.05, \delta'=0.1, \gamma=1.1, \psi=20, n=1000$)
  }
  \label{fig:energydist-pk}
\end{figure*}
\subsection{Generating high-quality seeds for further optimization}
Next, we empirically evaluated the capacity of \ourprog{} to generate seed sequences for multiple (pseudoknotted) target structures, possibly followed by subsequent local optimizations. As a baseline for comparison, we considered \RNAblueprint~\citep{Hammer2017}, the current leading tool for multiple design. 
As a quality measure, we used the multi-stable design objective function introduced by~\citet{Hammer2017}, defined as:
% \begin{align}
%   f(S)= \quad & \frac{1}{k} \sum_{\ell=1}^{k} (E(S, R_\ell) - G(S))\notag\\
%    & +\ \xi \frac{2}{k(k-1)} \sum\limits_{1\leq\ell<j\leq k}|E(S,R_\ell) - E(S,R_j)|.
% \end{align}
\begin{align}
  \label{eq:blueprintobjective}
    \Obj(S) &= \frac{1}{k} \sum_{\ell=1}^{k} (E(S, R_\ell) - \EnsE(S))\notag\\
   & +\frac{1}{2\binom{k}{2}} \sum\limits_{1\leq\ell<j\leq k}|E(S,R_\ell) - E(S,R_j)|.
\end{align}
where $\EnsE(S)$, the \Def{ensemble free energy} of $S$, is computed
by {\tt RNAfold}~\citep{Hofacker1994} in the pseudoknot-free case; for
pseudoknotted targets, $\EnsE(S)$ is approximated by the \Def{minimum
  free energy} (MFE) of S as estimated by {\tt HotKnots} in the energy model
of~\citet{dirks-pierce-03}.  Intuitively, the first term of \Obj{}
captures the distance of the targets from the ensemble free energy,
while the second term penalizes the dispersion of targets; \Obj{} is
best (minimal) when all targets simultaneously achieve the minimum
free energy of the sequence.

>>>>

%To reduce the cost of the precomputation, we used a stacking energy model
%as a proxy for the Turner model. Indeed, we observed a strong linear
%dependency between the Turner energy and the
%emphasize that only the simple stacking energies are directly targeted
%(at $\pm$10\% tolerance, except for a few hard instances). Due to our
%linear adjustment, we finally achieve well-controlled energies in the
%realistic Turner energy model.
\begin{figure*}[t]
  \begin{center}
    %{\sf \bfseries A}\
    \includegraphicstop[width=0.99\textwidth]{Figs/energy_shift.pdf}\hfill
    %{\sf \bfseries B}\ \includegraphicstop[width=0.55\textwidth]{Figs/PKB00211_PKB00239_0_energy_distribution}
  \end{center}
  \caption{%
    Targeting specific energies for pseudoknotted structures using
    multi-dimensional Boltzmann sampling.  A) Linear fits between the
    energies in the stacking model to the realistic pseudoknot energy
    model by Dirks and Pierce (D\&P) for sampled sequences; the good
    match enables more efficient targeting of Turner energies based on
    targeting stacking model energies.  We show the fits for both
    target structures S1 and S2 of subfigure B.  B) D\&P energy
    distributions for two target structures S1 and S2. We show the
    results of targeting the respective free energies $(-30,-20)$,
    $(-30,-30)$, $(-25,-25)$, and $(-20,-30)$ for the target
    structures (annotated in red, yellow, and blue
    resepectively)---this demonstrates the effectivity of our adaptive
    multi-dimensional Boltzmann sampling procedure; moreover, for
    comparison, the distributions for uniform and Boltzmann
    samples---respectively associated with homogenous weights $1$ and
    $e^\beta$.
%    ($\delta=0.05, \delta'=0.1, \gamma=1.1, \psi=20, n=1000$)
  }
  \label{fig:energydist-pk}
\end{figure*}
\subsection{Generating high-quality seeds for further optimization}
Next, we empirically evaluated the capacity of \ourprog{} to generate seed sequences for multiple (pseudoknotted) target structures, possibly followed by subsequent local optimizations. As a baseline for comparison, we considered \RNAblueprint~\citep{Hammer2017}, the current leading tool for multiple design.
As a quality measure, we used the multi-stable design objective function introduced by~\citet{Hammer2017}, defined as:
% \begin{align}
%   f(S)= \quad & \frac{1}{k} \sum_{\ell=1}^{k} (E(S, R_\ell) - G(S))\notag\\
%    & +\ \xi \frac{2}{k(k-1)} \sum\limits_{1\leq\ell<j\leq k}|E(S,R_\ell) - E(S,R_j)|.
% \end{align}
\begin{align}
  \label{eq:blueprintobjective}
    \Obj(S) &= \frac{1}{k} \sum_{\ell=1}^{k} (E(S, R_\ell) - \EnsE(S))\notag\\
   & +\frac{1}{2\binom{k}{2}} \sum\limits_{1\leq\ell<j\leq k}|E(S,R_\ell) - E(S,R_j)|.
\end{align}

where $\EnsE(S)$, the \Def{ensemble free energy} of $S$, is computed
by {\tt RNAfold}~\citep{Lorenz2011} in the pseudoknot-free case; for
pseudoknotted targets, $\EnsE(S)$ is approximated by the \Def{minimum
  free energy} (MFE) of S as estimated by {\tt HotKnots} in the energy model
of~\citet{dirks-pierce-03}.  Intuitively, the first term of \Obj{}
captures the distance of the targets from the ensemble free energy,
while the second term penalizes the dispersion of targets; \Obj{} is
best (minimal) when all targets simultaneously achieve the minimum
free energy of the sequence.
